---
title: "case_study_two"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Call Libraries
```{r}
library(zoo)
library(varian)
library(tidyverse)
library(corrplot)
library(glmnet)
library(FactoMineR)
library(factoextra)
library(MASS)
library(caret)
```

## Import Data

```{r import-file-names}
file_list <- list.files()
subject_data = file_list[grepl("df_S", file_list, fixed = TRUE)]
```

```{r}
# features_all_modalities = function(fn, win_size) {
#   df = read.csv(fn)
#   df = df[,-c("Label", "Subject")]
#   features_df <- data.frame(matrix(ncol = ncol(df)*6, nrow = nrow(df)))
#   new_names = sapply(1:length(df), function(x) {
#     c(paste0(colnames(df[,c]),"_mean"),
#       paste0(colnames(df[,c]),"_sd"),
#       paste0(colnames(df[,c]),"_range"),
#       paste0(colnames(df[,c]),"_min"),
#       paste0(colnames(df[,c]),"_max"),
#       paste0(colnames(df[,c]),"_skew")
#       )
#   })
#   colnames(features_df) = new_names
#   for (r in 1:nrow(df)) {
#     window = df[r:r+win_size*4,]
#     for (c in 1:length(window) ) {
#       if (colnames(window[,c]) != "Label" & colnames(window[,c]) != "Subject") {
#            column = window[,c]
#            
#            mu_column = mean(column)
# 
#            new_col_name = paste0(colnames(window[,c]),"_mean")
#            cindx = colnames(features_df)[whichcolnames(features_df)==new_col_name]
#            features_df[r, cindx] = mu_column
#       }
#     }
#  #make sure merge label subject back into feature dataframe
#   }
#   i
# }
```


```{r calc-stats-features}
# features_all_modalities = function(fn, win_size, shift) {
#   df_all = read.csv(fn) %>%
#     mutate(ACC_chest_3D = sqrt(ACC_chest_X^2+ACC_chest_Y^2+ACC_chest_Z^2)) %>%
#     mutate(ACC_wrist_3D = sqrt(ACC_wrist_x^2+ACC_wrist_y^2+ACC_wrist_z^2))
#   drops <- c("X", "Label", "subject")
#   df = df_all[ , !(names(df_all) %in% drops)]
#   replace_rows =  length(rollapply(df[,1], width = win_size*4, by = shift, FUN = mean, align = "left"))
#   features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
#   new_names = sapply(1:length(df), function(c) {
#     c(paste0(colnames(df)[c],"_mean"),
#       paste0(colnames(df)[c],"_sd"),
#       #paste0(colnames(df)[c],"_range"),
#       paste0(colnames(df)[c],"_min"),
#       paste0(colnames(df)[c],"_max")
#       # paste0(colnames(df)[c],"_skew")
#       )
#   })
#   colnames(features_df) = new_names
#   
# 
#   
#   for (c in 1:length(df)) {
#     
#     # finding mu
#     mu_vals = rollapply(df[,c], width = win_size*4, by = shift, FUN = mean, align = "left")
#     new_col_name = paste0(colnames(df)[c],"_mean")
#     cindx = which(colnames(features_df)==new_col_name)
#     features_df[, cindx] = mu_vals
#     
#     # finding sd
#     sd_vals = rollapply(df[,c], width = win_size*4, by = shift, FUN = var, align = "left")
#     new_col_name = paste0(colnames(df)[c],"_sd")
#     cindx = which(colnames(features_df)==new_col_name)
#     features_df[, cindx] = sqrt(sd_vals)
#     
#     # finding max
#     max_vals = rollapply(df[,c], width = win_size*4, by = shift, FUN = max, align = "left")
#     new_col_name = paste0(colnames(df)[c],"_max")
#     cindx = which(colnames(features_df)==new_col_name)
#     features_df[, cindx] = max_vals
#     
#     # finding min
#     min_vals = rollapply(df[,c], width = win_size*4, by = shift, FUN = min, align = "left")
#     new_col_name = paste0(colnames(df)[c],"_min")
#     cindx = which(colnames(features_df)==new_col_name)
#     features_df[, cindx] = min_vals
#     
#   }
#   
#  # make sure merge label, subject back into feature dataframe
#   features_df$Label = df_all[1:replace_rows, "Label"]
#   features_df$Subject = df_all[1:replace_rows, "subject"]
#   return(features_df)
# }
```


```{r IBI-to-hrv}
#Chest (Non ACC) for each person
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S CHEST INFORMATION

chest_nonacc_features = function(fn) {
  df_all = read.csv(fn) 
  drops <- c("X", "Label", "subject")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 28*5, by = 7, FUN = mean, align = "left"))
  chest_nonacc_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(chest_nonacc_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c], width = 28*5, by = 7, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(chest_nonacc_features_df)==new_col_name)
    chest_nonacc_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 28*5, by = 7, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(chest_nonacc_features_df)==new_col_name)
    chest_nonacc_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 28*5, by = 7, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(chest_nonacc_features_df)==new_col_name)
    chest_nonacc_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 28*5, by = 7, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(chest_nonacc_features_df)==new_col_name)
    chest_nonacc_features_df[, cindx] = min_vals
    
  }
  
 # make sure merge label, subject back into feature dataframe
  chest_nonacc_features_df$Label = df_all[1:replace_rows, "Label"]
  chest_nonacc_features_df$Subject = df_all[1:replace_rows, "subject"]
  chest_nonacc_features_df$ID <- seq.int(nrow(chest_nonacc_features_df))

  return(chest_nonacc_features_df)
}
```


```{r}
#Chest ACC for each person
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S CHEST ACC INFORMATION

chest_acc_features = function(fn) {
  df_all = read.csv(fn) %>%
    mutate(ACC_chest_3D = sqrt(ACC_chest_X^2+ACC_chest_Y^2+ACC_chest_Z^2)) 
  drops <- c("X")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 28*60, by = 7, FUN = mean, align = "left"))
  chest_acc_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(chest_acc_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c], width = 28*60, by = 7, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(chest_acc_features_df)==new_col_name)
    chest_acc_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 28*60, by = 7, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(chest_acc_features_df)==new_col_name)
    chest_acc_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 28*60, by = 7, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(chest_acc_features_df)==new_col_name)
    chest_acc_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 28*60, by = 7, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(chest_acc_features_df)==new_col_name)
    chest_acc_features_df[, cindx] = min_vals
    
  }
  chest_acc_features_df$ID <- seq.int(nrow(chest_acc_features_df))
  return(chest_acc_features_df)
}
```

```{r}
#Wrist BVP 
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S BVP Data Frame

wrist_bvp_features = function(fn) {
  df_all = read.csv(fn) 
  drops <- c("X")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 64*5, by = 16, FUN = mean, align = "left"))
  wrist_bvp_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(wrist_bvp_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c], width = 64*5, by = 16, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(wrist_bvp_features_df)==new_col_name)
    wrist_bvp_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 64*5, by = 16, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(wrist_bvp_features_df)==new_col_name)
    wrist_bvp_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 64*5, by = 16, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(wrist_bvp_features_df)==new_col_name)
    wrist_bvp_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 64*5, by = 16, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(wrist_bvp_features_df)==new_col_name)
    wrist_bvp_features_df[, cindx] = min_vals
    
  }
  
 # make sure merge label, subject back into feature dataframe
  wrist_bvp_features_df$ID <- seq.int(nrow(wrist_bvp_features_df))
  return(wrist_bvp_features_df)
}

```

```{r}
#Wrist EDA 
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S EDA Data Frame

wrist_eda_features = function(fn) {
  df_all = read.csv(fn) 
  drops <- c("X")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 4*5, by = 1, FUN = mean, align = "left"))
  wrist_eda_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(wrist_eda_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c], width = 4*5, by = 1, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(wrist_eda_features_df)==new_col_name)
    wrist_eda_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 4*5, by = 1, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(wrist_eda_features_df)==new_col_name)
    wrist_eda_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 4*5, by = 1, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(wrist_eda_features_df)==new_col_name)
    wrist_eda_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 4*5, by = 1, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(wrist_eda_features_df)==new_col_name)
    wrist_eda_features_df[, cindx] = min_vals
    
  }
  
 # make sure merge label, subject back into feature dataframe
  wrist_eda_features_df$ID <- seq.int(nrow(wrist_eda_features_df))

  return(wrist_eda_features_df)
}

```

```{r}
#Wrist TEMP 
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S TEMP Data Frame

wrist_temp_features_df = function(fn) {
  df_all = read.csv(fn) 
  drops <- c("X")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 4*5, by = 1, FUN = mean, align = "left"))
  wrist_temp_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(wrist_temp_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c], width  = 4*5, by = 1, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(wrist_temp_features_df)==new_col_name)
    wrist_temp_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 4*5, by = 1, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(wrist_temp_features_df)==new_col_name)
    wrist_temp_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 4*5, by = 1, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(wrist_temp_features_df)==new_col_name)
    wrist_temp_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 4*5, by = 1, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(wrist_temp_features_df)==new_col_name)
    wrist_temp_features_df[, cindx] = min_vals
    
  }
  
 # make sure merge label, subject back into feature dataframe
  wrist_temp_features_df$ID <- seq.int(nrow(wrist_temp_features_df))
  return(wrist_temp_features_df)
}

```

```{r}
#WRIST ACC 
#FIRST WE NEED TO DEFINE LIST OF EACH PERSON'S ACC WRIST Data Frame

wrist_acc_features = function(fn) {
  df_all = read.csv(fn) %>%
  mutate(ACC_wrist_3D = sqrt(ACC_wrist_x^2+ACC_wrist_y^2+ACC_wrist_z^2))
  drops <- c("X")
  df = df_all[ , !(names(df_all) %in% drops)]
  replace_rows =  length(rollapply(df[,1],  width = 32*60, by = 8, FUN = mean, align = "left"))
  wrist_acc_features_df <- data.frame(matrix(ncol = ncol(df)*4, nrow = replace_rows))
  new_names = sapply(1:length(df), function(c) {
    c(paste0(colnames(df)[c],"_mean"),
      paste0(colnames(df)[c],"_sd"),
      paste0(colnames(df)[c],"_min"),
      paste0(colnames(df)[c],"_max")
      )
  })
  colnames(wrist_acc_features_df) = new_names

  for (c in 1:length(df)) {
    
    # finding mu
    mu_vals = rollapply(df[,c],width = 32*60, by = 8, FUN = mean, align = "left")
    new_col_name = paste0(colnames(df)[c],"_mean")
    cindx = which(colnames(wrist_acc_features_df)==new_col_name)
    wrist_acc_features_df[, cindx] = mu_vals
    
    # finding sd
    sd_vals = rollapply(df[,c],  width = 32*60, by = 8, FUN = var, align = "left")
    new_col_name = paste0(colnames(df)[c],"_sd")
    cindx = which(colnames(wrist_acc_features_df)==new_col_name)
    wrist_acc_features_df[, cindx] = sqrt(sd_vals)
    
    # finding max
    max_vals = rollapply(df[,c],  width = 32*60, by = 8, FUN = max, align = "left")
    new_col_name = paste0(colnames(df)[c],"_max")
    cindx = which(colnames(wrist_acc_features_df)==new_col_name)
    wrist_acc_features_df[, cindx] = max_vals
    
    # finding min
    min_vals = rollapply(df[,c], width = 32*60, by = 8, FUN = min, align = "left")
    new_col_name = paste0(colnames(df)[c],"_min")
    cindx = which(colnames(wrist_acc_features_df)==new_col_name)
    wrist_acc_features_df[, cindx] = min_vals
    
  }
  
 # make sure merge label, subject back into feature dataframe
  wrist_acc_features_df$ID <- seq.int(nrow(wrist_acc_features_df))
  return(wrist_acc_features_df)
}

```

```{r}
subjects <- c("S2", "S3", "S4", "S5", "S6", "S7", "S8", "S9","S10" ,"S11","S13", "S14","S15","S16","S17")
```

```{r calc-stats-features}
features_all_modalities = function(subject) {
  #first, find all filenames
  chest_nonacc_file = paste0("df_", subject, "_chest_no_acc.csv")
  chest_acc_file = paste0("df_", subject, "_chest_acc.csv")
  wrist_bvp_file = paste0("df_", subject, "_bvp_wrist.csv")
  wrist_eda_file = paste0("df_", subject, "_eda_wrist.csv")
  wrist_temp_file = paste0("df_", subject, "_temp_wrist.csv")
  wrist_acc_file = paste0("df_", subject, "_wrist_acc.csv")
  
  #then, call each function
  chest_nonacc = chest_nonacc_features(chest_nonacc_file)
  chest_acc = chest_acc_features(chest_acc_file)
  wrist_bvp = wrist_bvp_features(wrist_bvp_file)
  wrist_eda = wrist_eda_features(wrist_eda_file)
  wrist_temp = wrist_temp_features(wrist_temp_file)
  wrist_acc =  wrist_acc_features(wrist_acc_file)
  
  #then merge everything by ID
  all_chest =  merge(chest_nonacc, chest_acc, by="ID")
  chest_bvp = merge(all_chest, wrist_bvp, by="ID")
  chest_bvp_eda = merge(chest_bvp, wrist_eda, by="ID")
  chest_bvp_eda_temp = merge(chest_bvp_eda, wrist_temp, by="ID")
  all = merge(chest_bvp_eda_temp, wrist_acc, by = "ID")
  return(all)
}
```

```{r}
ALL_df = NULL
for (subject in subjects) {
  print(subject)
  feature_data = features_all_modalities(subject) #CHANGE FILE PATH 
  #subject_no = gsub("\\..*","", sub('.*_', '', subject_data[i]))
  HR_fn = paste0("~/case-study-2/WESAD/", subject, "/", subject, "_E4_Data/HR.csv") 
  
  #CHANGE FILEPATH to be : paste0("/hpc/group/sta440-f20/WESAD/WESAD/", subject_no, "/" )
  HR_data = read.csv(HR_fn)[-1,]
  hr = HR_calc(as.data.frame(HR_data), 5, 1)
  
  df_hr = data.frame(matrix(unlist(hr), nrow= length(hr$mu),
                            ncol=4, byrow = F))
  colnames(df_hr) = c("hr_wrist_mu","hr_wrist_sd", "hr_wrist_min", "hr_wrist_max")#, "hr_wrist_range")
  df_hr$ID =seq.int(nrow(df_hr))
  
  feature_data = feature_data[-c(1:40),]
  feature_data$ID <- seq.int(nrow(feature_data))
  
  S_df = merge(feature_data, df_hr, by="ID")
  ALL_df = rbind(ALL_df, S_df)
  
  ALL_df <- ALL_df %>%
    filter(Label == 2 |Label == 3)
}

write.csv(ALL_df, "master2_df.csv")
```

```{r}
# IBI_data_3 = read.csv("~/case-study-2/WESAD/S3/S3_E4_Data/IBI.csv")
# 
# RMSSD_calc = function(df, peaks) {
#   # for (i in 1:nrow(df)) {
#   #   if (i <= nrow(df)-peaks)
#   vals = rollapply(df[,2], width = peaks, by = 1, FUN = rmssd, align = "left")
#   #}
#   
#   return(vals)
# }
# 
# hrv = RMSSD_calc(IBI_data_3, 3)
# IBI_data_3[1:length(hrv),3] = hrv
# colnames(IBI_data_3)[3] = "HRV"
# IBI_data_3 = na.omit(IBI_data_3)
```

```{r hrv-to-hz}
# hz_format = function(df, win_size) {
#   mu_vals = NULL
#   sd_vals = NULL
#   
#   for (i in 1:nrow(df)) {
#     #if (i <= nrow(df)-win_size)
#     nexttime = df[i,1] + win_size
#       # browser()
#     
#     #print(nexttime)
#     if (i+1 <= nrow(df)) {
# 
#     for (j in i+1:nrow(df)) {
#            # print(df[j,1])
#       if (!is.na(df[j,1])) {
#       #   print(df[j,1]==nexttime)
#       # if (df[j,1]==nexttime) {
#       #   stoprow = j
#       # }
#         # print(i)
#         # print(j)
#         # print(df[j,1])
#         # print(nexttime)
#       if ((df[j,1] >= nexttime) & (df[j-1,1]<nexttime)) {
#         stoprow = j-1
#       }
#       }
#     }
#     mu_vals = c(mu_vals, mean(df[i:stoprow, 3]))
#     sd_vals = c(sd_vals, sqrt(var(df[i:stoprow, 3])))
#     
#     }
#   }
#   
#   return(list(hrv_mu = mu_vals, hrv_sd = sd_vals))
# }

# test = hz_format(IBI_data_3, 5)
```

```{r hr-calc}
HR_calc = function(df, win_size, shift) {
  df <- df[rep(seq_len(nrow(df)), each = 4), ]
  mu_vals = rollapply(df, width = win_size*4, by = shift, FUN = mean, align = "left")
  sd_vals = rollapply(df, width = win_size*4, by = shift, FUN = sd, align = "left")
  min_vals = rollapply(df, width = win_size*4, by = shift, FUN = min, align = "left")
  max_vals = rollapply(df, width = win_size*4, by = shift, FUN = max, align = "left")

  return(list(mu = mu_vals, sd = sd_vals,
              min = min_vals, max = max_vals))
}
```

```{r run-all-featengine-combine }
ALL_df = NULL
for (i in 1:length(subject_data)) {
  feature_data = features_all_modalities(subject_data[i], 5, 1) #CHANGE FILE PATH 
  subject_no = gsub("\\..*","", sub('.*_', '', subject_data[i]))
  HR_fn = paste0("~/case-study-2/WESAD/", subject_no, "/", subject_no, "_E4_Data/HR.csv") #CHANGE FILEPATH to be : paste0("/hpc/group/sta440-f20/WESAD/WESAD/", subject_no, "/" )
  HR_data = read.csv(HR_fn)[-1,]
  hr = HR_calc(as.data.frame(HR_data), 5, 1)

  df_hr = data.frame(matrix(unlist(hr), nrow= length(hr$mu),
                             ncol=4, byrow = F))
  colnames(df_hr) = c("hr_wrist_mu","hr_wrist_sd", "hr_wrist_min", "hr_wrist_max")#, "hr_wrist_range")
  df_hr$ID =seq.int(nrow(df_hr))
  
  feature_data = feature_data[-c(1:40),]
  feature_data$ID <- seq.int(nrow(feature_data))

  S_df = merge(feature_data, df_hr, by="ID")
  ALL_df = rbind(ALL_df, S_df)
}
ALL_df = ALL_df %>% filter(Label %in% c("2","3"))
write.csv(ALL_df, "master_df.csv")
```

```{r corrplot}
ALL_df = read.csv("master_df.csv", header = T)[,-c(1:2)]

# get rid of chest EDA and temp and all min max (underdispersion)
remove_select_cov = ALL_df %>%
  dplyr::select(-contains("ACC_wrist")) %>%
  dplyr::select(-EDA_mean, -EDA_sd, -EDA_max, -EDA_min) %>%
  dplyr::select(-Temp_mean, -Temp_sd, -Temp_max, -Temp_min) %>%
  dplyr::select(-contains("BVP")) %>%
  dplyr::select(-contains("ECG")) %>%
  dplyr::select(-contains("min")) %>%
  dplyr::select(-contains("max")) %>%
  dplyr::select(-contains("ACC_chest_X")) %>%
  dplyr::select(-contains("ACC_chest_Y")) %>%
  dplyr::select(-contains("ACC_chest_Z"))

remove_select_cov$Label <- ifelse(remove_select_cov$Label=="2", 0, 1)

non_biological = c("Label", "Subject")
remove_select_cov_nb = remove_select_cov[ , !(names(remove_select_cov) %in% non_biological)]

# maybe reorder the variables
cormatrix = cor(remove_select_cov_nb)
corrplot(cormatrix, method = "color", addCoef.col="black", tl.cex = 0.4, number.cex= 10/ncol(remove_select_cov_nb))
```

# PCA

```{r pca}
scaled_select_cov_nb = scale(remove_select_cov_nb, center = TRUE, scale = TRUE)

# scaled_select_cov_nb %>% scaled_select_cov_nb
#   filter()
  
res.pca <- PCA(scaled_select_cov_nb, graph = FALSE, ncp = 5)
eig.val <- get_eigenvalue(res.pca)
eig.val
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```


```{r pca-cont}
var <- get_pca_var(res.pca)
corrplot(var$cos2, is.corr=FALSE, tl.cex = 0.5, number.cex= 3/ncol(remove_select_cov_nb))

fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 4, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 5, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 6, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 7, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 8, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 9, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 10, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 11, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 12, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 13, top = 10)
```
Dim1: very obviously ACC 
Dim2: EDA and temp + some ACC
Dim3: ECG BVP
Dim4: ECG EDA
Dim5: HR


starting from dim 8 from a domain perspective sutff doesn't matter

PCA basically confirms what we know about the correlations among the variables: we see two components that contains a lot of the ACC variables and another component that combines temperature and electrodermal activity (which makes sense because conditions such as sweating is correlated with temperature). It also makes sense that ECG and respiratory variables would be put with hear rate variables into a component.

```{r}
pca_vals <- as.data.frame(res.pca$ind$coord)
#test = as.data.frame(scale(res.pca$ind$coord, center = TRUE, scale=FALSE))
pca_vals$Label = remove_select_cov$Label
pca_vals$Subject = remove_select_cov$Subject
```

```{r }
smp_size <- floor(0.75 * nrow(pca_vals))
train_ind <- sample(seq_len(nrow(pca_vals)), size = smp_size)
train_pca <- pca_vals[train_ind, ]
test_pca <- pca_vals[-train_ind, ]



# FULL PCA MODEL model with main effect for Subject but no interactions (there is warning)
pca.model = glm(Label ~ ., family = binomial(link="logit"), data=train_pca)

pca.model.nsubj = glm(Label ~ .-Subject, family = binomial(link="logit"), data=train_pca)


# HOPEFULLY THE RIGHT ONE model with no main effect for Subject but with interactions
pca.model.interact = glm(Label ~ Dim.1+Dim.2+Dim.3+Dim.4+Dim.5+Dim.1:Subject
                  +Dim.2:Subject
                  +Dim.3:Subject
                  +Dim.4:Subject
                  +Dim.5:Subject
                    , family = binomial(link="logit"), data=train_pca)
summary(pca.model)

pred.test = pca.model %>% predict(test_pca, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(test_pca$Label==pred.test)
```

```{r}
remove_select_cov_nsubj = remove_select_cov %>% dplyr::select(-Subject, -Label)
scaled_remove_select_cov = scale(remove_select_cov_nsubj, center=TRUE, scale = TRUE)
scaled_remove_select_cov = cbind(scaled_remove_select_cov, as.data.frame(remove_select_cov$Subject), as.data.frame(remove_select_cov$Label))
scaled_remove_select_cov = scaled_remove_select_cov %>%
  rename(Subject = `remove_select_cov$Subject`) %>%
  rename(Label = `remove_select_cov$Label`)

train_raw <- scaled_remove_select_cov[train_ind, ]
test_raw <- scaled_remove_select_cov[-train_ind, ]

# OG FULL model with non-pca variables and PERF/QUASI separation
raw.model = glm(Label ~., family = binomial(link="logit"), data=train_raw)

raw.model = glm(Label ~. -Subject, family = binomial(link="logit"), data=train_raw)

# we have IDENTIFIABILITY issues 
summary(raw.model)

pred.test = raw.model %>% predict(test_raw, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(test_pca$Label==pred.test)

step.raw.model = raw.model %>% stepAIC(trace=FALSE)
summary(step.raw.model)
# this look sus

# GLMNET SUCKS
#convert training data to matrix format
train_raw_x = train_raw %>% 
  dplyr::select(-Label) %>%
  mutate(Subject = as.factor(Subject))

x <- data.matrix(train_raw_x)#, nrow = nrow(train_raw_x))
#convert class to numerical variable
y <- train_raw$Label
lasso.fit = glmnet(x, y, alpha = 1, family = "binomial")

plot(lasso.fit)

elasticnet.fit = glmnet(x, y, alpha = 0.5, family = "binomial")

plot(elasticnet.fit)
# print(lasso.fit)
# coefficients = coef(fit, s = 0.057370)

# cvfit <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure = "mse")
# plot(cvfit)

# min = coef(cvfit, s = "lambda.min")
# se1 = coef(cvfit, s = "lambda.1se")
```


```{r wristcorrplot}
ALL_df = read.csv("master_df.csv", header = T)[,-c(1:2)]

wrist_only = ALL_df %>%
  dplyr::select(contains("wrist")) %>%
  dplyr::select(-contains("ACC_wrist_x")) %>%
  dplyr::select(-contains("ACC_wrist_y")) %>%
  dplyr::select(-contains("ACC_wrist_z"))

wrist_only = cbind(wrist_only, as.data.frame(ALL_df$Subject), as.data.frame(ALL_df$Label))
wrist_only = wrist_only %>%
  rename(Subject = `ALL_df$Subject`) %>%
  rename(Label = `ALL_df$Label`) %>%
  mutate(Label = ifelse(Label==2, 0, 1)) %>%
  dplyr::select(-contains("min")) %>%
  dplyr::select(-contains("max")) 

# get rid of chest EDA and temp and all min max (underdispersion)
# remove_select_cov = ALL_df %>%
#   select(-contains("ACC_wrist")) %>%
#   select(-EDA_mean, -EDA_sd, -EDA_max, -EDA_min) %>%
#   select(-Temp_mean, -Temp_sd, -Temp_max, -Temp_min) %>%
#   select(-contains("BVP")) %>%
#   select(-contains("ECG")) %>%
#   select(-contains("min")) %>%
#   select(-contains("max")) %>%


# remove_select_cov$Label <- ifelse(remove_select_cov$Label=="2", 0, 1)

non_biological = c("Label", "Subject")
wrist_nb = wrist_only[ , !(names(wrist_only) %in% non_biological)]

# maybe reorder the variables
cormatrix = cor(wrist_nb)
corrplot(cormatrix, method = "color", addCoef.col="black", tl.cex = 0.6, number.cex= 10/ncol(wrist_nb))
```

# wrist PCA

```{r pca}
scaled_wrist_nb = scale(wrist_nb, center = TRUE, scale = TRUE)

# scaled_select_cov_nb %>% scaled_select_cov_nb
#   filter()
  
res.pca <- PCA(scaled_wrist_nb, graph = FALSE, ncp = 5)
eig.val <- get_eigenvalue(res.pca)
eig.val
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```


```{r pca-cont}
var <- get_pca_var(res.pca)
corrplot(var$cos2, is.corr=FALSE, tl.cex = 0.5, number.cex= 3/ncol(remove_select_cov_nb))

fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 4, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 5, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 6, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 7, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 8, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 9, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 10, top = 10)
fviz_contrib(res.pca, choice = "var", axes = 11, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 12, top = 10)
# fviz_contrib(res.pca, choice = "var", axes = 13, top = 10)
```

```{r run-wrist-model}
wrist_only_nsubj = wrist_only %>% dplyr::select(-Subject, -Label)
scaled_wrist_only = scale(wrist_only_nsubj, center=TRUE, scale = FALSE)
scaled_wrist_only = cbind(scaled_wrist_only, as.data.frame(wrist_only$Subject), as.data.frame(wrist_only$Label))
scaled_wrist_only = scaled_wrist_only %>%
  rename(Subject = `wrist_only$Subject`) %>%
  rename(Label = `wrist_only$Label`)

smp_size <- floor(0.75 * nrow(scaled_wrist_only))
train_ind <- sample(seq_len(nrow(scaled_wrist_only)), size = smp_size)
train_wrist <- scaled_wrist_only[train_ind, ]
test_wrist <- scaled_wrist_only[-train_ind, ]

# OUR MODEL FOR WRIST
wrist.full = glm(Label ~. -Subject, family = "binomial", data = train_wrist)
summary(wrist.full)

pred.test = wrist.full %>% predict(test_wrist, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(test_wrist$Label==pred.test)

step.wrist.model = wrist.full %>% stepAIC(trace=FALSE)
summary(step.wrist.model)

#convert training data to matrix format
train_wrist_x = train_wrist %>% 
  dplyr::select(-Label) %>%
  mutate(Subject = as.factor(Subject))

x <- data.matrix(train_wrist_x)#, nrow = nrow(train_raw_x))
#convert class to numerical variable
y <- train_wrist$Label
lasso.fit = glmnet(x, y, alpha = 1, family = "binomial")

plot(lasso.fit)

elasticnet.fit = glmnet(x, y, alpha = 0.5, family = "binomial")

plot(elasticnet.fit)
```

```{r}
# model one: combo data, pca, no subject
# model two: wrist data, no pca, no subject

# model three (interpret): combo data (hopefully), subject interactions no main effects
```

```{r kfold-cv}
set.seed(123)

kfoldcv_wrist_full = function(df) {
  
  actual_values = c()
  predicted_values = c()
  fitted_probs = c()
  
  # fold_accuracies = c()
  # fold_f1 = c()
  
  fold_no = c()
  subject = c()
  
  print("at folds")
  folds <- createFolds(factor(df$Subject), k = 5)
  
  print("past folds")
  for (i in 1:5) {
  #train and test
    col = paste0("Fold", i)
    indx = unlist(folds[col])

  train<- df[-indx,]
  print(nrow(train1))
  test<- df[indx, ]
  print(nrow(test1))
  
  
  # run model
  wrist.full = glm(Label ~. -Subject, family = "binomial", data = train)
#summary(wrist.full)

pred.test = wrist.full %>% predict(test, type= "response")
fitted_probs = c(fitted_probs, pred.test)
pred.test = ifelse(pred.test>0.5, 1,0)
predicted_values = c(predicted_values, pred.test)
actual_values = c(actual_values, test$Label)


  subject = c(subject, test$Subject)
  fold_no = c(fold_no, rep(i, nrow(test)))
  
  
  
  #acc1 = mean(test1$Label==pred.test)
  #fold_accuracies = c(fold_accuracies, acc1)
  
  # add function to compute f1
  # blah
}
  #train on training set
  #test on testing set 
  #predict labels
  #find accuracy and f1, add to fold_accuracies and fold_f1
  #put predicted values in predicted_values
  #put actual values in actual_values
  
  cvoutput = data.frame(cbind(actual_values, predicted_values, fitted_probs, 
        fold_no, subject))
  
  return(cvoutput)
}

wristcvoutput = kfoldcv_wrist_full(scaled_wrist_only)
```


```{r}
fold1 = wristcvoutput %>%
  filter(subject=="S11")#%>%
  #mean(actual_values==predicted_values)
mean(fold1$actual_values==fold1$predicted_value)
```

# SOME PLOTS MAY BE USEFUL
```{r distrib-vs-label}
varvslabel <- function(var){
  ggplot(remove_select_cov, aes_(y=as.name(var), x=remove_select_cov$Label, 
                                 group = remove_select_cov$Label)) +
    geom_boxplot() +
    labs(x ="Label", y = as.name(var))
}

lapply(names(remove_select_cov_nb), varvslabel)
```

It appears from these boxplots (prob most all except Electrodermal Activity to appendix), the distribution of physiological measures don't differ much between the stress and amusement states, but the boxplot for EDA shows a difference in distribution.


```{r subject-vs-covariates}
varvssubj <- function(var){
  ggplot(remove_select_cov, aes_(y=as.name(var), x=remove_select_cov$Subject, 
                                 group = remove_select_cov$Subject)) +
    geom_boxplot() +
    labs(x ="Subject", y = as.name(var))
}

lapply(names(remove_select_cov_nb), varvssubj)
```
For ACC, EDA, Temp, and HR variables, the mean, min, max data showed difference in distribution across the subjects. For ECG and BVP variables, only the sd measure showed differences across subjects.

# FROM SATURDAY

```{r testtrainsplit}
set.seed(123)
smp_size <- floor(0.75 * nrow(remove_select_cov))
train_ind <- sample(seq_len(nrow(remove_select_cov)), size = smp_size)
train <- remove_select_cov[train_ind, ]
# temp = train %>% select(-Label, -Subject)
# temp_centered = data.frame(scale(temp, center=TRUE, scale = FALSE)) 
# temp_centered$Label = train$Label
# train = temp_centered
test <- remove_select_cov[-train_ind, ]
# test = test %>% select(-Label, -Subject)
# temp_centered = data.frame(scale(temp, center=TRUE, scale = FALSE)) 
# temp_centered$Label = test$Label
# test = temp_centered

table(train$Label, train$Subject)
table(test$Label, test$Subject)
```

```{r trying-to-solve-separation}




  temp = remove_select_cov %>%
    filter(Subject == "S10")
  variability_plots = function(var) { ggplot(temp, aes_(y=as.name(var), x = temp$Label, group = temp$Label)) + geom_boxplot() + labs(x = "Label", y = as.name(var)) }
lapply(names(temp), variability_plots)
                 
   #bplist = c(bplist, plist)              

```

```{r glmnet-data-driven}
#convert training data to matrix format
x <- as.matrix(train%>%select(-Label), nrow = nrow(train))
#convert class to numerical variable
y <- ifelse(train$Label=="2", 0, 1)
fit = glmnet(x, y, alpha = 1, family = "binomial")

plot(fit)
print(fit)
coefficients = coef(fit, s = 0.057370)

cvfit <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure = "mse")
plot(cvfit)

min = coef(cvfit, s = "lambda.min")
se1 = coef(cvfit, s = "lambda.1se")
```

Now, let's use a domain driven approach.
We know stuff is correlated, so we don't want to use everything

Pick: acc_3D, EDA_min, EDA_max, temp_min, hr_mean, resp_sd
from a physiological standpoint, we know sweat (EDA), heating up (temp), breathing (resp), heart beating (hr),some justification for ACC makes it seem we should include those in some capcity in our model

```{r domain-driven}
set.seed(123)
domain.model = glm(Label ~ 1#+ #ACC_chest_3D_mean+ACC_chest_3D_sd#+ACC_chest_3D_min+ACC_chest_3D_max
                #   +EDA_wrist_mean+EDA_wrist_sd#+EDA_wrist_min+EDA_wrist_max
                #   +TEMP_wrist_mean+TEMP_wrist_sd#+TEMP_wrist_min+TEMP_wrist_max
                #   +Resp_mean+Resp_sd#+Resp_min+Resp_max
                #   +hr_wrist_mu+hr_wrist_sd#+hr_wrist_min+hr_wrist_max
                #   +Subject:ACC_chest_3D_mean+Subject:ACC_chest_3D_sd#+Subject:ACC_chest_3D_min+Subject:ACC_chest_3D_max
                   , family = binomial(link="logit"), data=train)

install.packages("lme4")
library(lme4)



fixed.domain.model = glmer(Label ~ #scale(ACC_chest_3D_mean)
                            # +scale(ACC_chest_3D_sd)#+ACC_chest_3D_min+ACC_chest_3D_max
                  # +scale(EDA_wrist_mean)#+scale(EDA_wrist_sd)#+EDA_wrist_min+EDA_wrist_max
                   #+scale(TEMP_wrist_mean)+scale(TEMP_wrist_sd)#+TEMP_wrist_min+TEMP_wrist_max
                   #+scale(Resp_mean)+scale(Resp_sd)#+Resp_min+Resp_max
                   #+scale(hr_wrist_mu)
                   #+scale(hr_wrist_sd)   +
                    (1|Subject)
     , data= train,family=binomial(link="logit"))

summary(fixed.domain.model)

pred.test = fixed.domain.model %>% predict(test, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(testlabels==pred.test)
#+Subject#+ACC_chest_3D_mean*Subject+
                     # ACC_chest_3D_mean*Subject+
                     # ACC_chest_3D_mean*Subject+
                     # ACC_chest_3D_mean*Subject+

#library(bestglm)
# domain.model = bestglm(Xy = train, family = binomial, IC = "AIC", RequireFullEnumerationQ = T)

# TRY BAS

summary(domain.model)

pred.test = domain.model %>% predict(test, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(testlabels==pred.test)

# step.domain.model = domain.model %>% stepAIC(trace=FALSE)
# summary(step.domain.model)
```

```{r}
# SUBJECT INCLUDED HERE
full.model = glm(Label ~., family = binomial(link="logit"), data=train)

pred.test = full.model %>% predict(test, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(test$Label==pred.test)

# SUBJECT ONLY MODEL
subj.model = glm(Label ~Subject, family = binomial(link="logit"), data=train)

pred.test = subj.model %>% predict(test, type= "response")
pred.test = ifelse(pred.test>0.5, 1,0)

mean(test$Label==pred.test)
```

```{r}
# IBI and HRV and HR

# WORK
# EDA
# # ggpairs
# # pca (Cathy)
# # clusters (Malavi)
# # plot by subject for interactions (Malavi)
# # 0/1 label boxplots versus (Cathy)

# models
# # frequentist Logistic regression w/ interactions
# # stepwise selection/diff selection methods (sensitivity analysis)
# # (DT, Randon Forest. Linear Discriminant Analysis, KNN)
# # for ML (put everything) use feature (variable) importance

# WORK
# # random forest (malavi)
# # Lda/knn (cathy)

# CROSS VALIDATION
# # amy said LOOCV will be computationally expensive
# # start with LOOCV (one person)
```
